---
title: "Homework 1"
author: "Amy Pitts"
date: "9/11/2020"
output: github_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(kableExtra)
#library(tibble)
```

## Problem 1
```{r Create_df}
prob1_df = 
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vec = factor(c("low", "low", "low", "mod", "mod", "mod", "high", "high", "high", "high"))
  )
prob1_df
```
We are now taking the mean for each variable 
```{r mean_var}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```
So far the random sample is the only one that I know should work. This is becuase they are numeric values. The vec_logic variable also produces results which I find interesting because it is actually a true and false value. I bet the mean is representing how often true comes up. The last to variables do not work and that is because they are not numeric values.

Number and logical and okay for taking the mean but not character. 

```{r try_numeric}
as.numeric(pull(prob1_df, samp))
as.numeric(pull(prob1_df, samp_gt_0)) #this works
as.numeric(pull(prob1_df, char_vec)) #this does not
as.numeric(pull(prob1_df, factor_vec)) #this one worked
```

The as numeric function works on the logic statement and the factored variable but not the character variable. The logic statement was assigned 0 and 1s while the factored variable was assigned 1 through 3. If i had factored the character variable before doing the as numeric function then the character variables would have been numbers rather than NA. 

```{r more_numeric}
#convert the logical vector to numeric, and multiply the random sample by the result
as.numeric(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)

#convert the logical vector to a factor, and multiply the random sample by the result
factor(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)

#convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result
as.numeric(factor(pull(prob1_df, samp_gt_0))) * pull(prob1_df, samp)
```


## Problem 2
```{r Prob2_df}
data("penguins", package = "palmerpenguins")
```

```{r explore_df}
dim(penguins) #printing th dimentions
variable.names(penguins) #the vraible names
mean(penguins$flipper_length_mm, na.rm=TRUE) #mean flipper length
```
This is a 8 variable dataset with 344 entries. The 8 variable names are listed above. The mean flipper length is 200.9152

```{r summary, warning = FALSE}
summary(penguins) #this gives us a summary of the varaibles. 
```
From this quick summary I see that species and island are factored with three options. Next the sex variable is also two options of female and male, with 11 missing values. The rest of the variables are numerical and there are missing values in each. Since the missing values will mess up our analysis we will want to ignore them for now. Since there are a low number this is okay, but if there was a bigger number of missing numbers we would have to worry about the introduction of bias. 

I am now going to summarize the data by grouping the pengiuns first by species and then by island. 
```{r more_explore, warning = FALSE, message = FALSE}
penguins %>% 
  group_by(species, island) %>%
  summarise(n = n(),
            bill_lenth_mean = mean(bill_length_mm, na.rm = TRUE),
            bill_dept_mean  = mean(bill_depth_mm, na.rm = TRUE), 
            flipper_lenth_mean = mean(flipper_length_mm, na.rm = TRUE),
            body_mass_mean = mean(body_mass_g, na.rm = TRUE)) 

```


First thing we notice is that Adelie penguin is on all three island while the Chinstrap is on the Dream island and the Gentoo only on the Biscoe island.  The next thing I notice is that the bill depth is much smaller and the body mass seems to be much bigger for the Gentoo pengiun compared to all the species. Overall the Adelie pengiun and the Chinstrap seem to be a little more comparable with the bill length being a little bigger than the Adelie comparing the means of the variables. 

We can now make a scatter plot of variables to see the relation. 
```{r plotting, warning = FALSE}
p <- ggplot(penguins, aes( bill_length_mm, flipper_length_mm)) + 
    geom_point(aes(colour = species)) + 
    ggtitle("Comparing Flipper Length and Bill Length between Three Penguin Species") +
    xlab("Bill Length (mm)") +
    ylab("Flipper Length (mm)")
p
#Exporing the graph and saving it in the porject directory. 
ggsave("scatter_plot_penguin.pdf", height = 4, width = 6) 
```



